---
title: "Lecture 5 Logistic, Poisson and Negative Binomial Regressions"
author: "Eryn McFarlane"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
  pdf_document: default
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pscl)
library(MASS)
```

We're going to simulate some predictor variables for each logistic regression, poisson regression and negative bionomial regression, and then analyze each. 

## Logistic Regression
- We're going to have one response variable and two predictor variables (one which is yes or no, the other which is continuous).

### QUESTION: What are you imagining your response variable to be? What are your two predictor variables?
- I am imagining a forced choice task (the most common psychophysics experiment methodology), where someone is trying to discern if the first stimuli or second stimuli is faster

```{r logistic regression}
N=50
predictor_cont<-rnorm(N,0,0.2) 
predictor_binary<-rbinom(N,1,0.5)

y<-1.44*predictor_cont+2.5*predictor_binary+rnorm(N, 0, 0.01) ### Remember that to make this do-able in a linear framework, like this, we're assuming this is on a log scale!
exp(1.44)

### We're now going to draw a probability distribution from y, assuming that y is on a logit scale
y_prob<-plogis(y)
y_prob

### we can now draw from the probability distribution (so that all of our responses are 0,1)

y_dummy = rbinom(n = N, size = 1, prob = y_prob)


data<-cbind.data.frame(1:N, y_prob, y, predictor_cont, predictor_binary)

log_reg_model<-glm(y_dummy~predictor_binary+predictor_cont, data=data, family='binomial')

### if you have loaded and installed lmerTest, it will give you p-values in your summary, if not, it won't
summary(log_reg_model)
### From here, you need to back transform to get the estimates on the y scale, instead of the log(y) scale

exp(coef(log_reg_model)) ### these are on the 'regular' scale, not the log scale
exp(confint(log_reg_model))
pR2(log_reg_model) ### r2CU (Nagelkerke R2 is the most comparable to our typical R2)



```
###Question, tell me about the distribution of y_prob?!
- It would be a distribution between 0-1 and mostly focus on 1, because my predictor_cont has a low SD

### Question: What did you expect the effect sizes of the predictors to be, based on what you simulated (hint, you will need to transform from line 25)?
- The effect sizes would be 1.44 and 2.5, but they wont be those exact values as a tiny bit of noise was added

### Question: The coefficient for predictor_cont is 1.44 This means that the expected log OR for a one-unit increase in predictor_cont is 1.44. This means that the OR (biological scale) for a one unit increase in predictor_cont is 4.220696. The R2 for this model is 0.2516362. This means that 25.1% of the variance can be explained by the model. 

Let's simulate some Poisson data!

## Poisson Regression

This is mostly borrowed from [Simulate Simulate Simulate 3!](https://aosmith.rbind.io/2018/07/18/simulate-poisson-edition/)

### Question: What is your imagined response variable here? What is your imagined predictor?
- Count of correct responses in the stimuli presented to a participant
### Question: What happens if you add more than one predictor (i.e. x2)?
- Then I am modelling 2 different effects on one outcome, as opposed to one

```{r Poisson Regression}

x1 = rnorm(1000,1,0.2) ### here there is one predictor, this a good place to imagine some biology

### Question, 

lambda = exp(0.5 + 2.5*x1) ### this is where I change the effect sizes for each predictor
exp(2.5)
###remember that this means that log(lambda) is a linear model!!
### here I'm simulating three betas against my three predictor variables to make my lambda
### The step above simulates the mean of each value of the response variable. lambda values are continuous, not discrete counts.
## λ is the unobserved true mean of the Poisson distribution for the t-th observation

head(lambda)

y = rpois(1000, lambda = lambda) ### error is added in this step, remembering that the mean and the variance are the same!
### remember that y is count data!
head(y)

plot(x1, log(y)) ### note the residuals - are they normally distributed? Is that okay?
plot(x1,y)

data_poisson<-cbind.data.frame(y, x1)

poisson_model<-glm(y ~ x1, family="poisson", data=data_poisson)
summary(poisson_model)

###back transform the coefficients 

exp(coef(poisson_model)) ### these are on the 'regular' scale, not the log scale
exp(confint(poisson_model))
pR2(poisson_model) ### r2CU (Nagelkerke R2 is the most comparable to our typical R2) 

```

### Question: The coefficient for x1 is 2.5 (FILL ME IN!) This means that the expected log count for a one-unit increase in x1 is 2.5. This means that the count difference (biological scale) for a one unit increase in x1 is 12.18249. The R2 for this model is 0.9956404. This means that the model explains 99.56% of the variation. 



```{r Negative Binomial Regression}
library(MASS)

neg_binom_predictor <- rnorm(1000, mean = 0, sd = 1) ### what is your imagined predictor?
# - my imagined predictor would be arm length 

# Log link function: log(mu) = beta0 + beta1 * X
mu <- exp(5+1.5 * neg_binom_predictor) ### notice that this is ALSO on a log scale
### mu is the expected mean for each observation
exp(1.5)
# Dispersion parameter for the negative binomial distribution
theta <- 2 ### this makes the difference between negative binomial and poisson. A very large theta would lead the negative binomial to approximate a poisson
##Play around with theta and see what happens!


y_neg_binom <- rnegbin(1000, mu = mu, theta = theta)

# Create a data frame to store the simulated data
data_negbiom <- cbind.data.frame(y_neg_binom, neg_binom_predictor)

plot(neg_binom_predictor, log(y_neg_binom))
plot(neg_binom_predictor, y_neg_binom)




negative_binomial_model<-glm.nb(y_neg_binom ~ neg_binom_predictor, data=data_negbiom)
summary(negative_binomial_model)

exp(coef(negative_binomial_model)) ### these are on the 'regular' scale, not the log scale
exp(confint(negative_binomial_model))
pR2(negative_binomial_model)

### r2CU (Nagelkerke R2 is the most comparable to our typical R2)
```

### Question: The coefficient for neg_binom_predictor is 1.5 (FILL ME IN!) This means that the expected log count for a one-unit increase in neg_binom_predictor is 1.5. This means that the count difference (biological scale) for a one unit increase in neg_binom_predictor is 4.481689. The R2 for this model is 0.8342348 This means that 83.42% of the model explains the variance seen in the data. 
