---
title: "Lecture 4 Multiple Regression"
author: "Ahmed Nadeem"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Multiple Regression

We're back to simulating the predictor variables directly, and then taking the response from those - this is a more typical way to simulate!

QUESTION: What parameters are you imagining in your simulations? - ANSWER: I am imagining a simulation that involves angles or error(distance in cm) from a target point. 

```{r Simulate Multiple parameters which affect the response variable}
library(MASS) ### this might be a package you need to install

##simulate multivariate normal data - this means that our parameters are related to each other
sigma <- matrix(c(1,   0.8,   0.5,
                  0.8,    0.7, 0.4,
                  0.5,0.4,   1), nrow = 3) ## positive-definite symmetric matrix specifying the covariance matrix of the variables - what does this mean? Make you sure can explain to yourself what this line does.

### This line gives us 3 parameters - how many do we want?
predictors<-mvrnorm(1000,rep(0,3), sigma) ###we are simulating the predictors here. What have we told this command to do? How correlated are the predictors?
cor.test(predictors[,1], predictors[,2])

```

QUESTION: How does changing sigma on line 22 change the correlations on line 28? - ANSWER: Making the numbers smaller makes the corelation go towards 1 (never hits 1 unless everything is 1)


QUESTION: What response variable are you imagining? How is it distributed? -> ANSAWER: the response variable i am thinking about (to continue from last week) is the error in position accuracy, it is normally distributed
QUESTION: What relationship are you simulating between your predictor variables and your response variable? -> i am simulating the strenght each cue has on generating an error
QUESTION: How much error are you simulating? How much irreduciable error do you expect in your response variable of interest? -> for the rNORM term, I am simulating A LOT OF ERROR, because humans are generally really bad when it comes to psycophysical tasks, but I also do not know anything about this reserach as it's relatively novel to measure the perception of space (at least according to my prof + any reading I've done).

```{r Simulate a response variable which all three variables affects the response variable}

response<-5.5*predictors[,1]+2.1*predictors[,2]+0.05*predictors[,3]+rnorm(1000, mean = 2, sd = 10) ### change this line based on what your response variable and imagined predictors are. Note that I have simulated a relationship between all predictors and the response, and note how much error I have simulated. 

cbind.data.frame(response, predictors)->data
names(data)<-c("cue_strength", "visual", "proprecptive", "vestibular") ### PUT INFORMATIVE NAMES HERE FOR WHAT YOU'RE SIMULATING
summary(data)
plot(data)

```

An assumption of multiple regression is that there isn't multicolinarity. 
We are breaking this assumption. 

QUESTION: Explore what happens to your R^2 when you change the strength of correlation between the paramters. - ANSWER: If I adjust the numbers above, the bigger I make them, the smaller the R^2 becomes for my anova and summary below


```{r Run a multiple regression}

lm(cue_strength~visual+proprecptive+vestibular, data=data)->model1
summary(model1)
anova(model1)

```


There are a couple of ways to plot multiple regressions. Some folks like 3D plots, so they can see the plan of the regression incorporating two predictor variables. This isn't my favourite, as it only works for two variables. 

Here, I'm going to plot using partial residuals - essentially, we're going to plot the relationship of one variable after we've accounted for the other two.


```{r plot multiple regression using partial residuals}

plot(data$visual, resid(lm(cue_strength~proprecptive+vestibular, data=data))) ### Do this for each of the three parameters of interest!
abline(model1)


```

QUESTION: Report the results of your multiple regression as you would in a results section of a paper. - ANSWER: The multiple regression analysis found that visual cues (𝛽= 5.05, p<0.001) and proprioceptive cues (𝛽= 2.97, p=0.023) significantly increased cue strength, while vestibular cues had no significant effect (  β=0.027, p=0.948). There were significant interactions between visual and vestibular cues ( β=−2.24, p=0.032) and between proprioceptive and vestibular cues ( β=2.71, p=0.034), indicating that the effects of visual and proprioceptive cues on cue strength depend on the level of vestibular input. The model explained 36.2% of the variance in cue strength ( 𝑅 2 = 0.3668 R 2 =0.3668).



```{r Run a multiple regression with interactions}

lm(cue_strength~visual*proprecptive*vestibular, data=data)->model2
summary(model2)
anova(model2)

```
QUESTION: Do you have evidence of an interaction? How does an interaction change the p-values or R^2? Why is this the case? - ANSWER: I think I have evidence, in the visual:vestibular interaction component. As well as the proprecptive:vestibular

An interaction is when the relationship between the response variable and a predictor variable changes depending on an other predictor variable.

QUESTION: Think of an interaction term! - ANSWER:visual:proprecptive and proprecptive:vestibular (based off significance)
```{r simulate an interaction}
predictor_4<-rnorm(1000,0,1)

response2<-1.5*predictors[,1]+ifelse(predictors[,1]>0.4, 0.5, 0)*predictor_4+rnorm(1000, 0, 1) ### make sure that you understand for yourself what this line does. Change this line so that it makes sense for your simulations!

cbind.data.frame(data, response2, predictor_4)->data2

names(data2)<-c("cue_strength", "visual", "propreceptive", "vestibular", "audio", "smell")
```

QUESTION: Describe in words what your interaction term is doing. - It helps describe the  interaction between visual and proprecptive cue, as well as the proprecptive and vestibular cue

Below, I am testing a 4 way interaction. Explain why this might or might not be appropriate for your own simulated data. 

```{r analyze with the interaction}

lm(cue_strength~visual*propreceptive*vestibular*audio, data=data2)->model3
summary(model3)
anova(model3)

```

QUESTION: Write a statistical methods paragraph and a statistical results paragraph. Make sure that everything in your statistical methods is reported in your results! - We conducted a multiple linear regression analysis to explore how three factors—visual, proprioceptive, and vestibular cues—affect cue strength. We included the individual effects of each factor as well as their combined effects (interactions between visual and proprioceptive, visual and vestibular, proprioceptive and vestibular, and a three-way interaction of all three). All factors were analyzed at the same time to determine their significance. We assessed the model’s accuracy by checking how well it explained the variance in cue strength, using R^2 and adjusted R^2 , and determined statistical significance with a threshold 𝑝 p-value of 0.05.
The model showed that visual and proprioceptive cues significantly increased cue strength, while vestibular cues had no significant effect. There were important interactions: as vestibular input increased, the effect of visual cues decreased, and the effect of proprioceptive cues increased. The model explained about 36.2% of the variation in cue strength.


